---
title: "Renal pager data"
author: "J Ritchie"
date: "09/12/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Git/Pager_analysis/")

#DEFINE PACKAGES
require(readxl)
require(dplyr)
require(tidyr)
require(readr)
require(lubridate)

All_messages <- read_excel("~/Dropbox/Pager data/outbox_all_messages%20James%20Ritchie.xlsx")

#Minor column tidy
All_messages$From<-NULL

#Select only renal pages
#set vector of renal numbers (note only one used here as EDA demonstrates all messages duplicated to both numbers)
Renal<-"'447623372487"

#select only these from the master DF
Renal_pages<-All_messages[All_messages$To %in% Renal,] 

#exclude failed pages
Renal_pages<-subset(Renal_pages, Renal_pages$Status!="Invalid Number")

#then drop the number identifier 
Renal_pages$To<-NULL

#drop initial DF for memory efficiency
rm(All_messages)
```

## Introduction

```{r, echo=FALSE}
Date_min<-min(Renal_pages$`Date Time`)
Date_min<-format(Date_min, '%d %B, %Y')

Date_max<-max(Renal_pages$`Date Time`)
Date_max<-format(Date_max, '%d %B, %Y')

Total_pages<-nrow(Renal_pages)

```


This is a summary document relating to pager activity for the renal baton.

The data do not account for calls to

-- Personal pagers / mobiles

-- Consultant on call

These data cover the period between `r Date_min` and `r Date_max` and describe `r Total_pages` pages sent out (note this number included multiple pages / reminders for individual encounters).

```{r , echo=FALSE}
Status_summary<-summary(Renal_pages$Status)
```

*****

The system utilises a two-way pager system for calls originated via swtichboard. Doctors can acknowledge calls to aid communication with waiting referrers. Overall acknowledgement is as follows:


```{r , echo=FALSE}

Status_table<-Renal_pages[,3]

table(Status_table)
```

This, representative of all calls including those from within SRFT


```{r, echo=FALSE, include=FALSE}

#Sepraate message into self contained units to allow extraction of numbers calling
Renal_pages<- Renal_pages %>% separate(Message, c("Time_value", "Part_2", "Part_3", "Part_4", "Part_5", "Part_6", "Part_7", "Part_8", "Part_9", "Part_10", "Part_11", "Part_12", "Part_13", "Part_14", "Part_15", "Part_16"), " ")

#Function to remove character values
  Text_remove<- function(x) {
    x<<-as.numeric(gsub("\\D", "", (x)))              
  }

Renal_pages[3:17]<-lapply(Renal_pages[3:17], Text_remove)

#Unite the multiple columns and then remove the NA data
Renal_pages<-unite(Renal_pages, "Number", c(3:17))

Renal_pages[3]<-Text_remove(Renal_pages$Number)

Renal_pages$Number<-as.numeric(Renal_pages$Number)


#NOW need to exclude internal numbers
#1 - Keep numbers with >5 digits as they represent full numbers e.g. 0161123456....
#2 - Keep numbers with 686 originator

#1 - 
Renal_pages$Full_number<-ifelse(Renal_pages$Number>70000,1,0)

#2 - 
Renal_pages$First_three<-substr(Renal_pages$Number, start=1, stop=3)
Renal_pages$Switch<-ifelse(Renal_pages$First_three=="686",1,0)

Renal_pages$External<-ifelse(Renal_pages$Full_number=="1"|Renal_pages$Switch=="1",1,0)

Renal_pages$Full_number<-NULL
Renal_pages$Switch<-NULL
Renal_pages$First_three<-NULL

#Reduce size of external numbers to aid data handling
#First remove incomplete cases
Renal_pages<-Renal_pages %>% drop_na()

Renal_pages$Number<-ifelse(Renal_pages$Number<70000,Renal_pages$Number, 12345)

#Create external call DF
External<-subset(Renal_pages, Renal_pages$External=="1")

External_n<-nrow(External)

Total_pages_Renal<-nrow(Renal_pages)

External_pct<-round(((External_n/Total_pages_Renal)*100), digits=1)
#Create internal call DF

Internal<-subset(Renal_pages, Renal_pages$External=="0")

Internal_n<-nrow(Internal)

Internal_pct<-round(((Internal_n/Total_pages_Renal)*100), digits=1)

```
Some pages did not have numeric data attached e.g. "Caller still waiting". As these cannot be linked to a specific episode they have been discarded

*****

The division of calls origins is:

-- Internal `r Internal_n` (`r Internal_pct`%)

-- External `r External_n` (`r External_pct`%)



Of the external pages, where the two-way functionality has value, responses were given as follows:

```{r , echo=FALSE}

External_status_table<-External[,4]

table(External_status_table)
```

## External calls

```{r, echo=FALSE}

#Now need to look at numbers in relation to time stamp to undertand how many multiple pages

#Split date / time into separate columns after duplication
External$Date_Time<-External$'Date Time'
External<- External %>% separate('Date Time', c("Date", "Time"), " ")
External$Time_value<-NULL

#Order calls by NUMBER then DATE_TIME

External<- External[order(External$Number, External$Date_Time),]

#to do - group by number, then mutate a new column based on the lag in timestamp
#Calculate the minutes difference between calls
External$mins <- unlist(tapply(External$Date_Time, INDEX = External$Number,FUN = function(x) c(0, `units<-`(diff(x), "mins"))))
#this does the whole column, resetting at number breaks - does not account for days etc - don't think this will influence the results.

#Have d/w switch; timeout break for calls being returned to them ins 15 minutes
External$Repeat_page<-ifelse(External$mins=="0"|External$mins>15,0,1)

Total_repeat<-sum(External$Repeat_page)

Repeat_pct<-round(((Total_repeat/(nrow(External)))*100), digits=1)
```

 Of the `r External_n` pages made in relation to external calls, `r Total_repeat` (`r Repeat_pct`%) were repeat calls due to the response taking > 15 mins
 
 
```{r, echo=FALSE}

External$Hour<-hour(External$Date_Time)
External$Hour<-as.numeric(External$Hour)

d <- density(External$Hour)
plot(d, main="Density plot of external referral timing")
polygon(d, col="red", border="blue")
```


## Externals - office hours
```{r, echo=FALSE}

Day_hours<-(9:17)

Daytime_external<-External[External$Hour %in% Day_hours,] 

d <- density(Daytime_external$Hour)
plot(d, main="Density plot of external referral timing (office hours)")
polygon(d, col="blue", border="red")

```



## Externals - on call hours
```{r, echo=FALSE}

Oncall_hours<-(c(1:8,18:23))

OOH_external<-External[External$Hour %in% Oncall_hours,] 

d <- density(OOH_external$Hour)
plot(d, main="Density plot of external call (on call hours)")
polygon(d, col="green", border="red")
```


##Externals - repeated calls

```{r, echo=FALSE}

Repeat_externals<-External[External$Repeat_page %in% 1,] 


#Repeat_externals<-subset(External, External$Repeat_page==1)

Repeat_externals_day<-Repeat_externals[Repeat_externals$Hour %in% Day_hours,] 


d <- density(Repeat_externals_day$Hour)
plot(d, main="Density plot of repeated external pages (office hours)")
polygon(d, col="yellow", border="blue")

```


## Internal calls

```{r, echo=FALSE}

#Now need to look at numbers in relation to time stamp to undertand how many multiple pages

#Split date / time into separate columns after duplication
Internal$Date_Time<-Internal$'Date Time'
Internal<- Internal %>% separate('Date Time', c("Date", "Time"), " ")
Internal$Time_value<-NULL

#Order calls by NUMBER then DATE_TIME

Internal<- Internal[order(Internal$Number, Internal$Date_Time),]

#to do - group by number, then mutate a new column based on the lag in timestamp
#Calculate the minutes difference between calls
Internal$mins <- unlist(tapply(Internal$Date_Time, INDEX = Internal$Number,FUN = function(x) c(0, `units<-`(diff(x), "mins"))))
#this does the whole column, resetting at number breaks - does not account for days etc - don't think this will influence the results.

#Have d/w switch; timeout break for calls being returned to them ins 15 minutes
Internal$Repeat_page<-ifelse(Internal$mins=="0"|Internal$mins>15,0,1)

Total_repeat_i<-sum(Internal$Repeat_page)

Repeat_pct_i<-round(((Total_repeat_i/(nrow(Internal)))*100), digits=1)
```

 Of the `r Internal_n` pages made in relation to external calls, `r Total_repeat_i` (`r Repeat_pct_i`%) were repeat calls due to the response taking > 15 mins
 
 
```{r, echo=FALSE}

Internal$Hour<-hour(Internal$Date_Time)
Internal$Hour<-as.numeric(Internal$Hour)

d <- density(Internal$Hour)
plot(d, main="Density plot of internal referral timing")
polygon(d, col="red", border="blue")
```

## Internals - office hours
```{r, echo=FALSE}

Day_hours<-(9:17)

Daytime_internal<-Internal[Internal$Hour %in% Day_hours,] 

d <- density(Daytime_internal$Hour)
plot(d, main="Density plot of internal referral timing (office hours)")
polygon(d, col="blue", border="red")

```

## Internals - on call hours
```{r, echo=FALSE}

Oncall_hours<-(c(1:8,18:23))

OOH_internal<-Internal[Internal$Hour %in% Oncall_hours,] 

d <- density(OOH_internal$Hour)
plot(d, main="Density plot of internal calls (on call hours)")
polygon(d, col="green", border="red")
```

## Internals - repeated calls
```{r, echo=FALSE}

Repeat_internals<-Internal[Internal$Repeat_page %in% 1,] 


#Repeat_externals<-subset(External, External$Repeat_page==1)

Repeat_internals_day<-Repeat_internals[Repeat_internals$Hour %in% Day_hours,] 


d <- density(Repeat_internals_day$Hour)
plot(d, main="Density plot of repeated internal pages (office hours)")
polygon(d, col="yellow", border="blue")

```

## Limitations

-- Assumptions about nature of repeated calls

-- Cannot account for other contemporaneous activiy

-- Consultant / personal calls excluded

-- Origin of calls not addressed

-- Data do not capture time to response

## Conclusions

-- Afternoon session has highest number of calls

-- Both internal and external calls are greatest in the afternoon

-- Need for additional support 1400-1600?

-- Excluding repeat calls (15 mins window), there were 2095 referrals to the baton

-- This is ~1050 calls per month

-- Assuming 30 calendar days this is 35 pages / day (higher than 12 months ago - had been 30). Activity may be increasing

-- No 'quiet time' for baton

-- Poor use of 2-ways functionality to communicate with switchboard

## Further work
-- Need to examine day to day variation (previous data show much lower number on Sunday)

-- Address how many were made around repeated call events (consider if non-response was likely due to other calls)

-- Consider system to allow time to response audit (likely JD driven)


